I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?

### 2023-12-08 output/naturalness-paper/paper.tex
updated paper

%\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i). It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past. 
%
%\begin{exe}
%\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
%\end{exe}
%}

Another issue with \citepos{mandelkern-etal2020} claim is that there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.%\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}

@@

\color{blue}

The linking function assumed in \citealt{mandelkern-etal2020} is that presupposition triggers are rated as unnatural in contexts in which the speaker is explicitly ignorant of the presupposed content, and that nontriggers are rated as natural in such contexts. This linking function is challenged by some of the results of our experiment. Consider first the CC of {\em be annoyed}, which received a very low mean naturalness rating in the explicit ignorance context. Under \citepos{mandelkern-etal2020} linking function, one can make sense of this result by analyzing the CC of {\em be annoyed} as being conventionally coded as having to be entailed by or satisfied in the common ground of the interlocutors. It has been long known, however, that such an analysis is not tenable for emotive predicates like {\em be annoyed}, as they do not presuppose that the CC is part of the common ground of the interlocutors but merely that the attitude holder believes the CC to be true (e.g., \citealt{heim92,karttunen2016}). In (\ref{emo}), for instance, the speaker does not believe that Sue could stay in bed.  \citealt{karttunen2016} suggested that the inference to speaker belief is a generalized conversational implicature that comes about in contexts that support the ``default assumption'' that the beliefs of the speaker and the attitude holder are aligned (p.712).

\begin{exe}
\ex\label{emo} 
Mary, who was under the illusion that it was Sunday, was glad that she could stay in bed. \\ \hspace*{.2cm} \hfill (\citealt{klein1975}, as cited in \citealt[122]{gazdar79a})
%Sally misremembered not having left a tip and regretted it. \hfill (\citealt[712]{karttunen2016})
\end{exe}
This analysis of {\em be annoyed} as merely presupposing attitude holder belief in the CC would, however, lead us to expect that {\em be annoyed} is rated as natural in contexts in which the speaker is explicitly ignorant of the CC. This is because the presupposition that the attitude holder believes that the CC is true does not conflict with the explicitly stated speaker ignorance about the truth of the CC. Why, then, did {\em be annoyed} receive such a low mean naturalness rating in explicit ignorance contexts?

\color{black}

One hypothesis is that naturalness ratings in explicit ignorance contexts are sensitive not just to whether the presupposition is defeasible (which is not the case for {\em too, also} and {\em again}) but also to at-issueness. On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content $c$ to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether $c$ is true (?$c$ e.g., {\em Does Julian dances salsa?}). The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?$c$. The naturalness of the interrogative is then modulated by whether the QUD ?$c$  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about $c$. As shown in \citealt{tbd-variability}, projective content differs in whether it can be at-issue. The CCs of emotive predicates, including {\em be annoyed}, are particularly resistant to being interpreted as at-issue. Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with a QUD about the CC (e.g., {\em Does Julian dance salsa?}). This incongruence may be why the CC of {\em be annoyed} received quite low naturalness ratings in the explicit ignorance context.

@@

\color{blue}

@@

\color{black}

@@

\subsection{Theoretical analyses of projection variability}

\color{blue}

As mentioned in \S\ref{s1}, \citealt{tbd-variability} proposed, based on the results of several experiments that used an inference rating measure, that projection is gradient.  \citealt{mandelkern-etal2020}, on the other hand, claimed that their naturalness rating measure categorically distinguishes presuppositions and nonpresuppositions.  However, as discussed above, neither the results presented in \citealt{mandelkern-etal2020} nor the results of our experiment empirically support the claim that projection is categorical and, instead, further supports the proposal that it is gradient. In other words, the empirical evidence that is available to date, and which is now based on both inference and naturalness rating measures, supports the proposal that by-content projection variability is gradient, not categorical. \jd{what does it mean for variability to be gradient? i thought the argument is that projection means form a gradient and there are different things that might mean theoretically. but i'm not sure what you have in mind when you say "by-content projection variability is gradient"}

This raises the question of how to analyze the observed projection variability. The analyses in \citealt{heim83} and \citealt{vds92} assume that expressions are either presupposition triggers or not, and they rely on local accommodation to account for examples in which presuppositions do not project.  As discussed in detail in \citealt{tbd-variability}, it is not clear that local accommodation can capture the observed by-content variation in a principled way, as analyses would need to assume that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}. Another problem such analyses face is that the CC of nonfactive predicates is also projective, sometimes even more so than the CC of factive predicates. Thus, as also discussed in \citealt{degen-tonhauser-language}, it is not clear which expressions would be analyzed as triggering projection, and which ones not.

Another set of projection analyses was developed to capture the projection of content associated with soft triggers (e.g., \citealt{abusch02,abusch10, abrusan2011,abrusan2016,romoli2015,simons-etal2017}). While these analyses differ in detail, they were developed to capture that the projection inferences associated with soft triggers are more easily suspended than those of hard triggers. These analyses predict s a three-way distinction in projection: Content that is conventionally marked as a presupposition (like the prejacent of {\em too} and {\em also}) is expected to be more projective than the content of soft triggers (like the prejacent of {\em stop} or the CC of {\em discover}), which in turn is expected to be more projective than content contributed by other expressions (like the CC of {\em think}). While this is promising, contemporary projection analyses do not suffice to account for observed by-content projection variability between soft triggers (e.g., that the CC of {\em know} is more projective than that of {\em discover}, which in turn is more projective than that of {\em reveal}), or that contents associated with nontriggers, such as the CCs of nonfactive predicates, may be projective, too, albeit to varying degrees (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion).

This situation suggests a promising avenue for future research that picks up a theme from \S\ref{s31}, namely to continue investigating distinctions between projection inferences. Understanding the dimensions along which projection inferences do or don't differ will lead to a refined taxonomy of such inferences (for a preliminary taxonomy see \citealt{brst-lang11}). The differences between empirically-motivated classes of projection inferences can, in turn, be decisive in identifying inferences already captured by extant analyses and inferences for which novel analyses must be developed.

\color{black}

>>>>

Another issue with \citepos{mandelkern-etal2020} claim is that there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and colleagues assumed that all presuppositions are unnatural in explicit ignorance contexts (modulo local accommodation), \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.

@@

The linking function assumed in \citealt{mandelkern-etal2020} is that presupposition triggers are rated as unnatural in contexts in which the speaker is explicitly ignorant of the presupposed content, and that nontriggers are rated as natural in such contexts. This linking function is challenged by some of the results of our experiment. Consider first the CC of {\em be annoyed}, which received a very low mean naturalness rating in the explicit ignorance context. Under \citepos{mandelkern-etal2020} linking function, one can make sense of this result by analyzing the CC of {\em be annoyed} as being conventionally coded as having to be entailed by or satisfied in the common ground of the interlocutors. It has been long known, however, that such an analysis is not tenable for emotive predicates like {\em be annoyed}, as they do not presuppose that the CC is part of the common ground of the interlocutors but merely that the attitude holder believes the CC to be true (e.g., \citealt{heim92,karttunen2016}). In (\ref{emo}), for instance, the speaker does not believe that Sue could stay in bed.

\begin{exe}
\ex\label{emo} 
Mary, who was under the illusion that it was Sunday, was glad that she could stay in bed. \\ \hspace*{.2cm} \hfill (\citealt{klein1975}, as cited in \citealt[122]{gazdar79a})
%Sally misremembered not having left a tip and regretted it. \hfill (\citealt[712]{karttunen2016})
\end{exe}
\citealt{karttunen2016} suggested that the inference to speaker belief is a generalized conversational implicature that comes about in contexts that support the ``default assumption'' that the beliefs of the speaker and the attitude holder are aligned (p.712). This analysis would, however, lead us to expect that {\em be annoyed} is rated as natural in contexts in explicit ignorance contexts because the presupposition that the attitude holder believes that the CC is true does not conflict with the explicitly stated speaker ignorance about the truth of the CC. Why, then, did {\em be annoyed} receive such a low mean naturalness rating in explicit ignorance contexts?

One hypothesis is that naturalness ratings in explicit ignorance contexts are sensitive not just to whether a content is a nondefeasibile presupposition (as may be the case with {\em too, also} and {\em again}) but also to the at-issueness of content. On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content $c$ to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether $c$ is true (?$c$ e.g., {\em Does Julian dances salsa?}). The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?$c$. The naturalness of the interrogative is then modulated by whether the QUD ?$c$  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about $c$. As shown in \citealt{tbd-variability}, projective contents differ in at-issueness. The CCs of emotive predicates, including {\em be annoyed}, are particularly resistant to being interpreted as at-issue. Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with a QUD about the CC (e.g., {\em Does Julian dance salsa?}). This incongruence may be why the CC of {\em be annoyed} received quite low naturalness ratings in the explicit ignorance context.

@@

@@

@@

\subsection{Towards a formal analysis of gradient projection}

As mentioned in \S\ref{s1}, \citealt{tbd-variability} observed that the CCs of factive predicates vary in how projective they are,  based on data collected with an inference rating measure. This observation led them to propose that projection is gradient rather than categorical. This proposal was further supported by the results of the experiments and analyses of existing datasets in \citealt{degen-tonhauser-language}, who found that the projection means of the CCs of factive and nonfactive predicates form a gradient rather than being categorically distinct (see Fig.~\ref{fig:dt1a}). \citealt{mandelkern-etal2020}, on the other hand, maintained that projection is categorical and  they claimed that  their naturalness rating measure categorically distinguishes presuppositions and nonpresuppositions.  As discussed above, neither the results presented in \citealt{mandelkern-etal2020} nor the results of our experiment empirically support the claim that projection is categorical. Instead, the results of our experiment provide further empirical support for the proposal that contents associated with different expressions (e.g., the CCs of {\em know, discover} and {\em confess}) do not either project or not project, but rather exhibit quite fine-grained variation in how projective they are.

There are already some suggestions in the contemporary literature on projection for how to deal with projection variation. The analyses of \citealt{heim83} and \citealt{vds92}, which assume that expressions are either presupposition triggers or not, rely on local accommodation to account for examples in which presuppositions do not project (for discussion see, e.g., \citealt{djaerv-bacovcin2020,mandelkern-etal2020}). However, as discussed in detail in \citealt{tbd-variability}, it is not clear that local accommodation can capture the observed by-content variation in a principled way, as such analyses would need to assume, for instance, that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}. Another challenge for these analyses is the observation that the CCs of nonfactive predicates are also projective,  sometimes even more so than the CCs of factive predicates. Because nonfactive predicates do not trigger a presupposition in analyses like \citealt{heim83} and \citealt{vds92}, these analyses make no predictions about the projection of the CCs of nonfactive predicates (see \citealt{degen-tonhauser-language} for further discussion).

Another set of projection analyses was developed to capture the projection of content associated with soft triggers (e.g., \citealt{abusch02,abusch10, abrusan2011,abrusan2016,romoli2015,simons-etal2017}). These analyses (which differ in detail)  were developed to capture that the projection of content associated with soft triggers is more easily suspended than that of hard triggers. Given such analyses, a three-way distinction in projection can be predicted: Content that is conventionally marked as a presupposition (like the prejacent of {\em too} and {\em also}) is expected to be more projective than the content of soft triggers (like the prejacent of {\em stop} or the CC of {\em discover}), which in turn is expected to be more projective than content contributed by expressions that have traditionally not been analyzed as presuppositions (like the CC of {\em confess}). While this is a promising step in the right direction, it obviously falls short of fully capturing the observed projection variability, for instance, that the CC of {\em know} is more projective than that of {\em discover}, which in turn is more projective than that of {\em reveal}, or that contents associated with nontriggers, such as the CCs of nonfactive predicates, may be projective, too, albeit to varying degrees (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion).

This situation suggests that a promising avenue for future research is to continue investigating distinctions between expressions associated with projection inferences (see \S\ref{s31}).  Understanding the dimensions along which projection inferences do or don't differ will lead to a refined taxonomy of such inferences (for a preliminary taxonomy see \citealt{brst-lang11}). The differences between empirically-motivated classes of projection inferences can, in turn, be decisive in identifying inferences already captured by extant analyses and inferences for which novel analyses must be developed.
